---
title: "Getting started with onedevr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with onedevr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

`onedevr` is an R client for the [OneDev](https://onedev.io) REST API. It provides:

- **High-level helpers** (`od_*` functions) for common workflows
- **Low-level escape hatch** (`od_request()`) for custom queries
- **Tibble-first results** for easy data manipulation
- **Connection flexibility** (environment variables or explicit objects)

This vignette covers basic setup and common tasks. For detailed workflows, see:

- `vignette("issues", package = "onedevr")` — Issue management
- `vignette("builds", package = "onedevr")` — Build queries and artifacts
- `vignette("pull-requests", package = "onedevr")` — PR workflows

## Key Concepts

**UI Numbers vs. Internal IDs:** onedevr functions use **UI numbers** (e.g., `#145`), matching what you see in the OneDev web UI. The package handles internal ID resolution automatically.

**Tibbles by Default:** List and query functions return tibbles for easy piping and data manipulation:

```{r}
library(onedevr)

issues <- od_query_issues(state = "Open")
issues %>%
  dplyr::filter(priority == "High") %>%
  dplyr::select(number, title, assignee)
```

# Installation

Install from GitHub:

```r
remotes::install_github("alexseymer/onedevR")
```

Verify installation:

```r
library(onedevr)
packageVersion("onedevr")  # Should be 0.5.1 or later
```

# Configuration

## Quick Start: Environment Variables

The simplest setup uses environment variables. Set them in `.Renviron` or via `Sys.setenv()`:

```r
Sys.setenv(
  ONEDEV_HOST = "https://onedev.example.com",
  ONEDEV_API_TOKEN = "your-token-here",
  ONEDEV_PROJECT_PATH = "mygroup/myproject"
)

# Now you can use onedevr functions directly
od_query_issues(state = "Open")
```

**Available Environment Variables:**

| Variable | Purpose | Example |
|----------|---------|---------|
| `ONEDEV_HOST` | OneDev server URL | `https://onedev.example.com` |
| `ONEDEV_API_TOKEN` | Bearer token (recommended) | `(token)` |
| `ONEDEV_USERNAME` | For Basic Auth | `myuser` |
| `ONEDEV_PASSWORD` | For Basic Auth | `mypass` |
| `ONEDEV_PROJECT_PATH` | Default project | `group/project` |
| `ONEDEV_PROJECT_ID` | Default project ID | `1` |

See `.env.example` in the repository for additional variables.

## Advanced: Explicit Connection Objects

For fine-grained control or multiple connections:

```r
# Create a connection
conn <- od_connection(
  host = "https://onedev.example.com",
  token = Sys.getenv("ONEDEV_API_TOKEN"),
  project_path = "group/project"
)

# Set as package default (optional)
od_set_connection(conn)

# Or pass to individual functions
issues <- od_query_issues(state = "Open", conn = conn)
```

## Troubleshooting Connection Issues

**Error: `401 Unauthorized`**
- Check token is valid: `curl -H "Authorization: Bearer $TOKEN" https://onedev.example.com/~api/issues`
- Verify `ONEDEV_HOST` has no trailing slash
- Try Basic Auth if token isn't working: `ONEDEV_USERNAME` + `ONEDEV_PASSWORD`

**Error: `404 Not Found` (issues/builds)**
- Verify project path: `od_get_project()` to check configured project
- For project-independent queries, omit `ONEDEV_PROJECT_PATH`

# Working with Issues

## Query and Inspect

```r
library(onedevr)

# Get all open issues (limited to 50 by default)
open <- od_query_issues(state = "Open", count = 50L)

# Get specific issue by UI number
issue <- od_get_issue(145)  # Issue #145

# See available issue fields
fields <- od_get_issue_fields()

# Get comments on an issue
comments <- od_get_issue_comments(145)
```

## Create and Update

```r
# Create new issue
new_issue <- od_create_issue(
  title = "Add user authentication",
  description = "Implement OAuth2 support",
  fields = list(Priority = "High", Type = "Feature")
)

issue_number <- new_issue$number  # e.g., 146

# Update title and description
od_issue_set_title(issue_number, "Add OAuth2 authentication")
od_issue_set_description(issue_number, "Enhanced description here")

# Set custom fields
od_issue_set_fields(issue_number, Priority = "Critical", Assignee = "alice")

# Change state
od_issue_transition_state(issue_number, "In Progress")
od_issue_transition_state(issue_number, "Closed")
```

## Iterations and Links

```r
# List sprints/iterations
iterations <- od_list_iterations()

# Add issue to iteration
od_add_issue_iterations(145, iteration_ids = iterations$id[1])

# Add comment
od_add_issue_comment(
  issue_number = 145,
  content = "Working on this now",
  reply_to = NA  # or comment ID for threading
)
```

# Working with Builds

## Query Builds

```r
# Query by status (use OneDev keywords)
successful <- od_query_builds(status = "successful", count = 10L)
failed <- od_query_builds(status = "failed", count = 10L)

# See query language
cat(od_get_query_description("build"))

# Get specific build
build <- od_get_build(123)

# Build parameters and logs
params <- od_get_build_params(123)
logs <- od_get_build_log(123)  # Returns log lines
```

## Artifacts and Results

```r
# List build artifacts
artifacts <- od_list_build_artifacts(123)

# Download artifact
od_download_artifact(123, "dist/app.zip", "app-123.zip")

# Inspect results
results <- od_get_build_results(123)
```

## Trigger Builds

```r
# Run a job
od_run_job("CI", branch = "main", parameters = list(VERBOSE = TRUE))

# Rebuild existing build
od_rebuild_job(123)

# Cancel running build
od_cancel_job(123)
```

# Working with Pull Requests

```r
# Query PRs
open_prs <- od_query_pull_requests(status = "open", count = 20L)

# Get specific PR
pr <- od_get_pull_request(5)

# Get comments and reviews
comments <- od_get_pull_request_comments(5)
reviews <- od_get_pull_request_reviews(5)

# Add comment
od_add_pull_request_comment(5, "Looks good, merging now")

# Approve
od_approve_pull_request(5, "Approved and verified in CI")
```

# Repository Operations

```r
# Get project info
project <- od_get_project()
clone_url <- od_get_project_clone_url()  # http, ssh URLs

# List branches
branches <- od_list_branches()

# Get branch info
main <- od_get_branch("main")

# Query commits
commits <- od_query_commits(path = "R/", count = 20L)

# Get file contents
readme <- od_get_file_text("main", "README.md")

# List packages
packs <- od_query_packages(count = 10L)
```

# Pagination

For large result sets, use `od_paginate()`:

```r
# Fetch all issues in batches
all_issues <- od_paginate(
  od_query_issues,
  state = "Open",
  page_size = 50L,     # Items per page
  max_pages = Inf      # Fetch all available
)

# Or limit total pages
recent_50 <- od_paginate(
  od_query_builds,
  status = "successful",
  page_size = 10L,
  max_pages = 5L
)
```

# Low-Level API Access

For queries not covered by high-level helpers, use `od_request()`:

```r
# Make custom REST calls
issue_id <- od_resolve_issue_id(145)
raw <- od_request("GET", paste0("/issues/", issue_id))

# POST example
new_branch <- od_request(
  method = "POST",
  endpoint = "/repository/branches",
  data = list(name = "feature/new", baseBranch = "main")
)

# See full endpoint reference at:
# https://your-onedev-host/~help/api
```

# Best Practices

1. **Use environment variables** for shared scripts and CI/CD
2. **Cache results** when fetching large datasets
3. **Handle errors** gracefully — network timeouts can occur
4. **Verify queries** using `od_get_query_description()` for syntax
5. **Use pagination** for production scripts, not loop + offset
6. **Log operations** when making state changes (issue creation, build triggering)

# Examples and Recipes

See `vignette("issues", package = "onedevr")` for detailed workflows on:
- Bulk issue operations
- Custom queries
- Filtering and reporting

Additional resources:
- **OneDev API docs**: `https://your-host/~help/api`
- **TheOneDev CLI**: https://github.com/theonedev/tod
- **gitlabr reference**: https://thinkr-open.github.io/gitlabr/

# Troubleshooting

**Q: How do I use UI numbers vs. internal IDs?**
A: Always use UI numbers (e.g., `145` for `#145`). onedevr resolves them automatically.

**Q: What's the tibble option for?**
A: Set `as_tibble = FALSE` to get plain lists instead of tibbles:
```r
od_query_issues(as_tibble = FALSE)
# Or globally:
options(onedevr.as_tibble = FALSE)
```

**Q: Can I use this for multiple OneDev instances?**
A: Yes, create separate connection objects:
```r
prod <- od_connection(host = "https://prod.ondev", token = "...")
staging <- od_connection(host = "https://staging.ondev", token = "...")
od_query_issues(state = "Open", conn = prod)
```

**Q: Does onedevr support webhooks?**
A: Not directly. Use OneDev's webhook settings to POST to your R Shiny app or trigger CI.

---

For more help, file an issue: https://github.com/alexseymer/onedevR/issues
