---
title: "Getting started with onedevr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with onedevr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

`onedevr` is an R client for self-hosted [OneDev](https://onedev.io) REST APIs.
It follows the same shape as [`gitlabr`](https://thinkr-open.github.io/gitlabr/)
(`od_*` helpers + `od_request()` escape hatch) and mirrors OneDev conventions
from the [`tod`](https://github.com/theonedev/tod) CLI.

**Important:** high-level helpers take **UI numbers** (`145` / `"#145"`), not
internal REST ids.

List/query helpers return **tibbles** by default. Opt out with
`as_tibble = FALSE` or `options(onedevr.as_tibble = FALSE)`.

# Configure a connection

## Environment variables

```{r}
library(onedevr)

Sys.setenv(
  ONEDEV_HOST = "https://git.example.test",
  ONEDEV_API_TOKEN = "your-token",
  ONEDEV_PROJECT_PATH = "group/my-project"
)

od_query_issues(state = "Open")
```

See `.Renviron.example` / `.env.example` in the repository for the full
variable list (including Basic Auth via `ONEDEV_USERNAME` /
`ONEDEV_PASSWORD`).

## Explicit connection

```{r}
conn <- od_connection(
  host = "https://git.example.test",
  token = Sys.getenv("ONEDEV_API_TOKEN"),
  project_path = "group/my-project"
)

od_set_connection(conn) # optional package default
od_get_issue(145, conn = conn)
```

# Issues

```{r}
issues <- od_query_issues(state = "Open", count = 20L)
issue <- od_get_issue(145)
fields <- od_get_issue_fields(145)
comments <- od_get_issue_comments(145)

created <- od_create_issue(
  title = "API test",
  description = "Created from R",
  fields = list(Priority = "Normal")
)

od_issue_set_title(created$number, "API test (renamed)")
od_issue_transition_state(created$number, "Closed")
```

Iterations (sprint/version-like objects in OneDev):

```{r}
iterations <- od_list_iterations()
od_add_issue_iterations(145, iteration_ids = iterations$id[1])
```

# Builds, jobs, and artifacts

Build **status** filters use OneDev keyword criteria (`successful`, `failed`,
…), not a `"Status"` field. Enum spellings like `"SUCCESSFUL"` are accepted
and mapped.

```{r}
builds <- od_query_builds(status = "successful", count = 10L)
build <- od_get_build(100)
params <- od_get_build_params(100)
log_lines <- od_get_build_log(100)
artifacts <- od_list_build_artifacts(100)

# Run / rebuild / cancel (state-changing — use deliberately)
# od_run_job("CI", branch = "main")
# od_rebuild_job(100)
# od_cancel_job(100)
```

Inspect the server’s build query grammar:

```{r}
cat(od_get_query_description("build"))
```

# Pull requests

```{r}
open_prs <- od_query_pull_requests(status = "open", count = 10L)
pr <- od_get_pull_request(1)
comments <- od_get_pull_request_comments(1)
reviews <- od_get_pull_request_reviews(1)
```

# Projects, users, packages, and git

```{r}
projects <- od_query_projects(count = 20L)
project <- od_get_project()
clone <- od_get_project_clone_url() # fields: http, ssh

branches <- od_list_branches()
tip <- od_get_branch("main")
commits <- od_query_commits(count = 10L)
readme <- od_get_file_text("main", "README.md")

packs <- od_query_packages(count = 10L)
```

# Pagination

```{r}
all_open <- od_paginate(
  od_query_issues,
  state = "Open",
  page_size = 50L,
  max_pages = 20L
)
```

# Low-level escape hatch

```{r}
issue_id <- od_resolve_issue_id(145)
raw <- od_request("GET", paste0("/issues/", issue_id))
```
